---
title: "derived traits"
author: "Frederick J. Boehm"
date: "1/6/2020"
output: html_document
params: 
  lastmod: !r lubridate::now()
---

Last modified: `r params$lastmod`.


```{r setup2, include=FALSE, dev="CairoPNG"}
knitr::opts_chunk$set(dev="CairoPNG")
```


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

First, we read the derived traits from Clare's csv file.

```{r readcsv}
tr <- readr::read_csv("../data/CC_clust_tomap.csv")
```


Now, we load Rick's mapping objects. 

```{r load-ricks-mapping-stuff}
load("../data/reduced_map_qtl2_mapping_objects.Rdata")
```

We calculate leave-one-chromosome-out kinship matrices.

```{r}
kinship <- qtl2::calc_kinship(probs = genoprobs.1, type = "loco")
```



Next, run `scan1` from R package `qtl2`. 

```{r scan1}
tr_mat <- as.matrix(tr[, - c(1:3)])
rownames(tr_mat) <- tr$CC.Line
s1 <- qtl2::scan1(genoprobs = genoprobs.1, pheno = tr_mat, kinship = kinship, reml = TRUE)
```

Note that our univariate QTL mapping was done with linear mixed effects models.

```{r find_peaks}
library(magrittr)
qtl2::find_peaks(s1, map = map.1, threshold = 6) %>%
  dplyr::arrange(desc(lod))
```

Note that the top 3 peaks are all on Chr 10 and fairly close to each other.


```{r get-effects}
s1c_25 <- qtl2::scan1coef(genoprobs = genoprobs.1[,10], pheno = tr_mat[, 25], kinship = kinship$`10`, reml = TRUE)
s1c_23 <- qtl2::scan1coef(genoprobs = genoprobs.1[,10], pheno = tr_mat[, 23], kinship = kinship$`10`, reml = TRUE)
s1c_7 <- qtl2::scan1coef(genoprobs = genoprobs.1[,10], pheno = tr_mat[, 7], kinship = kinship$`10`, reml = TRUE)
```

```{r plots}
qtl2::plot_coefCC(s1c_25, map = map.1, scan1_output = s1[, 25, drop = FALSE])
qtl2::plot_coefCC(s1c_23, map = map.1, scan1_output = s1[, 23, drop = FALSE])
qtl2::plot_coefCC(s1c_7, map = map.1, scan1_output = s1[, 7, drop = FALSE])
```

We notice that all three plots have yellow and purple at opposite extremes.


## Permutations

Let's find permutation thresholds for the traits.

```{r perms}
fn <- "../data/derived-traits-perms-1000.rds"

if (!file.exists(fn)) {
  perms <- qtl2::scan1perm(genoprobs = genoprobs.1, 
                           pheno = tr_mat, 
                           kinship = kinship, 
                           n_perm = 1000, 
                           reml = TRUE, 
                           cores = 0 # for using multiple cores on the ghpcc cluster
                           )
  saveRDS(perms, fn)
} else {
  perms <- readRDS(fn)
}

```

```{r thresholds}
apply(X = perms, FUN = quantile, probs = 0.95, MARGIN = 2)
```

```{r get-pvalue}
qtl2::find_peaks(s1, map = map.1, threshold = 6) %>%
  dplyr::arrange(desc(lod)) %>%
  dplyr::mutate(pvalue = purrr::map2_dbl(.f = function(lod, lodindex, perms){qtl2effects::calc_pvalue(observed_lod = lod, maxlods = perms[ , lodindex])}, .x = lod, .y = lodindex, perms = perms)) %>%
  dplyr::arrange(pvalue)
```





## Other ideas

One might also consider multivariate QTL mapping in addition to mapping these "derived" traits one at a time.
