---
title: "Refining mega-hotspots"
author: "Frederick J. Boehm"
date: "12/18/2019"
output: html_document
params:
  lastmod: !r lubridate::now()
---

Last modified: `r params$lastmod`.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Rick previously calculated bayes intervals for the ten hotspots. Two of them, namely hotspots 4 and 6, on Chromosomes 10 and 11, are what we're calling "mega-hotspots" because they have many associated traits.

Rick's bayes intervals, per "hotspot_allele_effects5.html", for hotspots 4 and 6 are:

Hotspot 4: (0, 66)

Hotspot 6: (0, 73)


We now want to see if we can use the strongest peaks and their bayes intervals to refine our definition of each hotspot.

`qtl2` has a function, `bayes_int`, that uses a `scan1` output object and a map to determine a bayes interval for the peak.

Rick has stored the Neto hotspot results in "neto_traits_by_probe3_annotated.csv".

Let's read it into R. Recall that it has one row per marker (ie, marker within a hotspot).

```{r, read-neto}
library(magrittr)
neto <- readr::read_csv("../data/neto_traits_by_probe3_annotated.csv") %>%
  dplyr::filter(hs %in% c(4, 6)) %>%
  dplyr::select(- probe, - chr, - cM, - neto.n, - row, - n.traits) %>%
  tidyr::pivot_longer(cols = V2:V35) %>%
  dplyr::filter(!is.na(value))
```
We now have a collection of traits for which we must run `scan1`. 

```{r, filter}
neto_unique_traits <- neto %>%
  dplyr::filter(!duplicated(value))
```

We see that there are 62 unique traits between the two megahotspots.

Now, let's run `scan1` for the 62 traits.

```{r, prep-scan1}
# read the tnseq traits
tnseq <- readRDS("../data/tnseq-traits.rds")
hot_traits <- tnseq[, colnames(tnseq) %in% neto_unique_traits$value]
# load rick's mapping objects
load("../data/reduced_map_qtl2_mapping_objects.Rdata")
```

```{r, scan1}
kinship <- qtl2::calc_kinship(genoprobs.1, "loco")
fn <- "../data/scan1_hs46.rds"
if (!file.exists(fn)){
  s1hot <- qtl2::scan1(genoprobs = genoprobs.1, 
            pheno = hot_traits, 
            kinship = kinship, 
            addcovar = addcovar, 
            reml = TRUE, 
            cores = 0
            )
} else {
  s1hot <- readRDS(fn)
}
```

```{r findpeaks}
(p10 <- qtl2::find_peaks(s1hot, map = map.1, threshold = 6) %>%
  dplyr::filter(chr == 10) %>%
  dplyr::filter(lod == max(lod))
)
```


```{r bayes_intervals}
qtl2::bayes_int(scan1_output = s1hot[, p10$lodindex, drop = FALSE], map = map.1, chr = 10)
```


```{r findpeaks11}
(p11 <- qtl2::find_peaks(s1hot, map = map.1, threshold = 6) %>%
  dplyr::filter(chr == 11) %>%
  dplyr::filter(lod == max(lod))
)
```

```{r bayes_intervals11}
qtl2::bayes_int(scan1_output = s1hot[, p11$lodindex, drop = FALSE], map = map.1, chr = 11)
```

We'll look at the LODs for the Chromosome 11 hotspot.

```{r plotlod}
qtl2::plot_scan1(s1hot, map = map.1, lodcolumn = p11$lodindex, chr = 11)
```
The above analysis uses only the strongest peak for each mega-hotspot when determining the boundaries (ie, bayes credible intervals).

## Top traits in hotspot 4

```{r findpeaks-4}
thr <- 5
(p10_top <- qtl2::find_peaks(s1hot, map = map.1, threshold = thr) %>%
  dplyr::filter(chr == 10) %>%
  dplyr::filter(pos > 50) %>%
  dplyr::arrange(desc(lod))
)
```

Because I used the `find_peaks` function, I'm seeing only three traits with peaks at positions greater than 50 cM. Other traits might have high lods in this interval, but not have peaks in that interval. They thus wouldn't be in the table above.

Also, I used a threshold of `r thr` when defining the peaks. Some traits may have peaks below `r thr`.

Let's look at allele effects for the top three traits.

## Allele effects in hotspot 4

### Trait 53, hotspot 4

```{r, tr53}
# calc allele effects
qtl2::scan1coef(genoprobs = genoprobs.1[, 10], 
            pheno = hot_traits[, 53, drop = FALSE], 
            kinship = kinship[[10]], 
            addcovar = addcovar, 
            reml = TRUE) %>%
  qtl2::plot_coefCC(map = map.1[10], scan1_output = s1hot[, 53, drop = FALSE])
```

### Trait 48, hotspot 4


```{r, tr48}
# calc allele effects
qtl2::scan1coef(genoprobs = genoprobs.1[, 10], 
            pheno = hot_traits[, 48, drop = FALSE], 
            kinship = kinship[[10]], 
            addcovar = addcovar, 
            reml = TRUE) %>%
  qtl2::plot_coefCC(map = map.1[10], scan1_output = s1hot[, 48, drop = FALSE])
```

### Trait 57, hotspot 4

```{r, tr57}
# calc allele effects
qtl2::scan1coef(genoprobs = genoprobs.1[, 10], 
            pheno = hot_traits[, 57, drop = FALSE], 
            kinship = kinship[[10]], 
            addcovar = addcovar, 
            reml = TRUE) %>%
  qtl2::plot_coefCC(map = map.1[10], scan1_output = s1hot[, 57, drop = FALSE])
```

### Trait 30, hotspot 4

```{r, tr30}
# calc allele effects
qtl2::scan1coef(genoprobs = genoprobs.1[, 10], 
            pheno = hot_traits[, 30, drop = FALSE], 
            kinship = kinship[[10]], 
            addcovar = addcovar, 
            reml = TRUE) %>%
  qtl2::plot_coefCC(map = map.1[10], scan1_output = s1hot[, 30, drop = FALSE])
```


### Trait 10, hotspot 4

```{r, tr10}
# calc allele effects
qtl2::scan1coef(genoprobs = genoprobs.1[, 10], 
            pheno = hot_traits[, 10, drop = FALSE], 
            kinship = kinship[[10]], 
            addcovar = addcovar, 
            reml = TRUE) %>%
  qtl2::plot_coefCC(map = map.1[10], scan1_output = s1hot[, 10, drop = FALSE])
```


## Top traits for Hotspot 6

```{r findpeaks-6}
(top <- qtl2::find_peaks(s1hot, map = map.1, threshold = thr) %>%
  dplyr::filter(chr == 11) %>%
   dplyr::filter(pos > 50) %>% # ensure that peaks are near Rick's estimated hotspot position
  dplyr::arrange(desc(lod))
)
```

### Trait 21, hotspot 6

```{r, tr21}
# calc allele effects
qtl2::scan1coef(genoprobs = genoprobs.1[, 11], 
            pheno = hot_traits[, 21, drop = FALSE], 
            kinship = kinship[[11]], 
            addcovar = addcovar, 
            reml = TRUE) %>%
  qtl2::plot_coefCC(map = map.1[11], scan1_output = s1hot[, 21, drop = FALSE])
```

### Trait 26, hotspot 6

```{r, tr26}
# calc allele effects
qtl2::scan1coef(genoprobs = genoprobs.1[, 11], 
            pheno = hot_traits[, 26, drop = FALSE], 
            kinship = kinship[[11]], 
            addcovar = addcovar, 
            reml = TRUE) %>%
  qtl2::plot_coefCC(map = map.1[11], scan1_output = s1hot[, 26, drop = FALSE])
```

### Trait 41, hotspot 6

```{r, tr41}
# calc allele effects
qtl2::scan1coef(genoprobs = genoprobs.1[, 11], 
            pheno = hot_traits[, 41, drop = FALSE], 
            kinship = kinship[[11]], 
            addcovar = addcovar, 
            reml = TRUE) %>%
  qtl2::plot_coefCC(map = map.1[11], scan1_output = s1hot[, 41, drop = FALSE])
```

### Trait 32, hotspot 6

```{r, tr32}
# calc allele effects
qtl2::scan1coef(genoprobs = genoprobs.1[, 11], 
            pheno = hot_traits[, 32, drop = FALSE], 
            kinship = kinship[[11]], 
            addcovar = addcovar, 
            reml = TRUE) %>%
  qtl2::plot_coefCC(map = map.1[11], scan1_output = s1hot[, 32, drop = FALSE])
```

### Trait 19, hotspot 6

```{r, tr19}
# calc allele effects
qtl2::scan1coef(genoprobs = genoprobs.1[, 11], 
            pheno = hot_traits[, 19, drop = FALSE], 
            kinship = kinship[[11]], 
            addcovar = addcovar, 
            reml = TRUE) %>%
  qtl2::plot_coefCC(map = map.1[11], scan1_output = s1hot[, 19, drop = FALSE])
```

## Bayes intervals for strongest traits (near Rick's estimated hotspot positions)

We now calculate the bayes intervals, one per hotspot, for the trait with the strongest (highest) LOD near Rick's estimated hotspot positions (per Rick's v5 html).

```{r bayes10}
qtl2::bayes_int(scan1_output = s1hot[, 53, drop = FALSE], map = map.1, chr = 10)
```

```{r bayes11}
qtl2::bayes_int(scan1_output = s1hot[, 21, drop = FALSE], map = map.1, chr = 11)
```
