---
title: "Refining mega-hotspots"
author: "Frederick J. Boehm"
date: "12/18/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Rick previously calculated bayes intervals for the ten hotspots. Two of them, namely hotspots 4 and 6, on Chromosomes 10 and 11, are what we're calling "mega-hotspots" because they have many associated traits.

Rick's bayes intervals, per "hotspot_allele_effects5.html", for hotspots 4 and 6 are:

Hotspot 4: (0, 66)

Hotspot 6: (0, 73)


We now want to see if we can use the strongest peaks and their bayes intervals to refine our definition of each hotspot.

`qtl2` has a function, `bayes_int`, that uses a `scan1` output object and a map to determine a bayes interval for the peak.

Rick has stored the Neto hotspot results in "neto_traits_by_probe3_annotated.csv".

Let's read it into R. Recall that it has one row per marker (ie, marker within a hotspot).

```{r, read-neto}
library(magrittr)
neto <- readr::read_csv("../data/neto_traits_by_probe3_annotated.csv") %>%
  dplyr::filter(hs %in% c(4, 6)) %>%
  dplyr::select(- probe, - chr, - cM, - neto.n, - row, - n.traits) %>%
  tidyr::pivot_longer(cols = V2:V35) %>%
  dplyr::filter(!is.na(value))
```
We now have a collection of traits for which we must run `scan1`. 

```{r, filter}
neto_unique_traits <- neto %>%
  dplyr::filter(!duplicated(value))
```

We see that there are 62 unique traits between the two megahotspots.

Now, let's run `scan1` for the 62 traits.

```{r, prep-scan1}
# read the tnseq traits
tnseq <- readRDS("../data/tnseq-traits.rds")
hot_traits <- tnseq[, colnames(tnseq) %in% neto_unique_traits$value]
# load rick's mapping objects
load("../data/reduced_map_qtl2_mapping_objects.Rdata")
```

```{r, scan1}
kinship <- qtl2::calc_kinship(genoprobs.1, "loco")
fn <- "../data/scan1_hs46.rds"
if (!file.exists(fn)){
  s1hot <- qtl2::scan1(genoprobs = genoprobs.1, 
            pheno = hot_traits, 
            kinship = kinship, 
            addcovar = addcovar, 
            reml = TRUE, 
            cores = 0
            )
} else {
  s1hot <- readRDS(fn)
}
```

```{r findpeaks}
(p10 <- qtl2::find_peaks(s1hot, map = map.1, threshold = 6) %>%
  dplyr::filter(chr == 10) %>%
  dplyr::filter(lod == max(lod))
)
```

```{r bayes_intervals}
qtl2::bayes_int(scan1_output = s1hot[, p10$lodindex, drop = FALSE], map = map.1, chr = 10)
```


```{r findpeaks11}
(p11 <- qtl2::find_peaks(s1hot, map = map.1, threshold = 6) %>%
  dplyr::filter(chr == 11) %>%
  dplyr::filter(lod == max(lod))
)
```

```{r bayes_intervals11}
qtl2::bayes_int(scan1_output = s1hot[, p11$lodindex, drop = FALSE], map = map.1, chr = 11)
```

The above analysis uses only the strongest peak for each mega-hotspot when determining the boundaries (ie, bayes credible intervals).


