---
title: "Analyze Clinical Covariates"
author: "REB"
date: "5/21/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

##### Object

1. To determine if including batch covariates in the IL10 scans actually ipmrove the fit.

2. To investigate why the TIMBR results using batch covariates yield much "worse" results than those obtained using batch-corrected phenotype values and no covariates.

##### Procedure

Load data.

```{r data}
load("~/Desktop/TIMBR/CC_cs_36state.Rdata")
load("~/Desktop/TIMBR/reduced_map_qtl2_mapping_objects.Rdata")
pheno <- read.csv("/Users/bakerr/Desktop/rqtl2_loco_v4_mirror/batch_corrected_phenos_summarized_05OCT2019_v4.csv", row.names=1)
# remove parents
pheno <- pheno[grep("CC0", row.names(pheno)), ]
# fix row.names
tmp <- row.names(pheno)
tmp <- sub(".GeniUnc", "cs", tmp)
tmp <- sub(".TauUnc", "cs", tmp)
tmp <- sub(".Unc", "cs", tmp)
row.names(pheno) <- tmp
pheno.raw <- read.csv("/Users/bakerr/Desktop/rqtl2_loco_v4_mirror/pheno_clinical_4qtl2_win2_05OCT2019_v4.csv", row.names=1)
reduced.probes <- do.call(`c`, lapply(genoprobs.1, function(x) dimnames(x)[[3]]))
reduced.probes <- as.vector(reduced.probes)
keep <- dimnames(model.probs)[[3]] %in% reduced.probes
model.probs.reduced <- model.probs[row.names(pheno), , keep]
map.Mb <- qtl2convert::map_df_to_list(GM.snps, chr="Chrom", pos="pos")
# limit phenotypes to spleenCFU/CEQ, IL10
pheno.raw <- pheno.raw[, c(1,2,13)]
pheno <- pheno[, c(5,6,20)]
colnames(pheno)[1:2] <- c("spleenCFU", "spleenCEQ")
```

scan1 -- use uncorrected phenotype values + batch covariates for QTL scans.

```{r map}
library(qtl2)
kinship <- calc_kinship(genoprobs.1, "loco")
out <- scan1(genoprobs = genoprobs.1, pheno = pheno.raw, kinship = kinship, addcovar=addcovar, Xcovar = Xcovar, cores = 2)
# trait maxLOD probes
pos <- c(bayes_int(out, map.Mb, chr=2, lodcolumn=1)[2], bayes_int(out, map.Mb, chr=2, lodcolumn=2)[2], bayes_int(out, map.Mb, chr=2, lodcolumn=3)[2])
probes <- sapply(pos, find_marker, map=map.Mb, chr=2)
print(data.frame(probes=probes, pos=pos, row.names=colnames(pheno)))
```

Fit QTL's and extract the batch coefficients.

```{r effects, fig.align='center'}
cfu <- pheno.raw[, 1]
names(cfu) <- row.names(pheno)
ceq <- pheno.raw[, 2]
names(ceq) <- row.names(pheno)
il10 <- pheno.raw[, 3]
names(il10) <- row.names(pheno)
eff <- list()
idx <- 1
for (i in probes[1:2]) {
  eff[[idx]] <- fit1(genoprobs.1[[2]][,,i], cfu, kinship[[2]], addcovar=addcovar)$coef
  eff[[idx+2]] <- fit1(genoprobs.1[[2]][,,i], ceq, kinship[[2]], addcovar=addcovar)$coef
  eff[[idx+4]] <- fit1(genoprobs.1[[2]][,,i], il10, kinship[[2]], addcovar=addcovar)$coef
  idx <- idx + 1
}
eff.df <- do.call(cbind, eff)
colnames(eff.df) <- c(paste("spleenCFU @", round(pos[1:2], digits=2)), paste("spleenCEQ @", round(pos[1:2], digits=2)), paste("IL10 @", round(pos[1:2], digits=2)))
```

Get coefficients used for batch correction and compare to the actual scan1 batch coefficients.

```{r barplot, fig.align='center'}
batch.cor <- list()
for (i in 1:3) {
  diff <- pheno.raw[, i] - pheno[, i]
  t <- table(diff)
  t <- t[order(t, decreasing=T)]
  batch.cor[[i]] <- as.numeric(names(t))[1:3]
}
df <- do.call(rbind, batch.cor)
row.names(df) <- paste("lm", colnames(pheno), sep=":")
colnames(df) <- paste0("block", 1:3)
tmp <- cbind(eff.df[9:10, ], t(df[, 2:3]))
tmp <- tmp[, c(1,2,7,3,4,8,5,6,9)]
par(mar=c(8, 4, 4, 2) + 0.1, xpd=T)
barplot(tmp, beside=T, las=2, cex.names=0.8, ylab="coefficient")
legend("top", c("batch2", "batch3"), fill=c("grey30", "grey80"), ncol=2, inset=-0.25, cex=0.9)
```

The batch coefficients from spleen CFU/CEQ are quite similar to those obtained from simple regression of the phenotype values. The scan1 coefficients for IL10 are smaller (in absolute value). Is the QTL fit actually improved by including the covariates?

Test fit with and without covariates. [The comparison is made at the maxLOD probe from the covariate scan. It's possible/likely the scan w/o covariates may produce a different maxLOD probe!]

```{r test}
df <- as.data.frame(genoprobs.1[[2]][, , probes[1]])
df$spleenCFU <- pheno.raw$spleenCFU
fit1 <- lm(spleenCFU ~ ., df)
df <- cbind(df, addcovar[grep("CC0", row.names(addcovar)), ])
fit2 <- lm(spleenCFU ~ ., df)
print(lmtest::lrtest(fit1, fit2))
df <- as.data.frame(genoprobs.1[[2]][, , probes[1]])
df$spleenCEQ <- pheno.raw$spleenCEQ
fit1 <- lm(spleenCEQ ~ ., df)
df <- cbind(df, addcovar[grep("CC0", row.names(addcovar)), ])
fit2 <- lm(spleenCEQ ~ ., df)
print(lmtest::lrtest(fit1, fit2))
df <- as.data.frame(genoprobs.1[[2]][, , probes[1]])
df$IL10 <- pheno.raw$IL10
fit1 <- lm(IL10 ~ ., df)
df <- cbind(df, addcovar[grep("CC0", row.names(addcovar)), ])
fit2 <- lm(IL10 ~ ., df)
print(lmtest::lrtest(fit1, fit2))
```

It appears the covariates improve the fit, at least at this probe.

Find the maxLOD probes for scans using raw phenotype values and no covariates.

```{r no_covar}
out2 <- scan1(genoprobs = genoprobs.1, pheno = pheno.raw, kinship = kinship, Xcovar = Xcovar, cores = 2)
pos2 <- c(bayes_int(out2, map.Mb, chr=2, lodcolumn=1)[2], bayes_int(out2, map.Mb, chr=2, lodcolumn=2)[2], bayes_int(out2, map.Mb, chr=2, lodcolumn=3)[2])
probes2 <- sapply(pos2, find_marker, map=map.Mb, chr=2)
print(data.frame(probes=probes2, pos=pos2, row.names=colnames(pheno)))
```

The spleenCFU maxLOD shifts appreciably, spleenCEQ somewhat, but IL10 maps to exactly the same probe.

Use scan1 batch coefficients for QTL fit at 174.2925 Mb to correct raw IL10 values. Run TIMBR on the corrected phenotype values.

```{r TIMBR_1, eval=FALSE}
library(TIMBR)
data(mcv.data)
prior.M1 <- list(model.type = "crp", # crp - Chinese Restaurant Process
  prior.alpha.type = "gamma",
  prior.alpha.shape = 1,
  prior.alpha.rate = 2.333415)
addcovar <- addcovar[grep("CC0", row.names(addcovar)), ]
# probes[1]
b2 <- eff.df[9, 5]
b3 <- eff.df[10, 5]
corr <- pheno.raw
corr[addcovar$block2 == 1, "IL10"] <- corr[addcovar$block2 == 1, "IL10"] - b2
corr[addcovar$block3 == 1, "IL10"] <- corr[addcovar$block3 == 1, "IL10"] - b3
trait <- "IL10"
res <- list()
P <- model.probs.reduced[, , probes[1]]
prior.D <- mcv.data$prior.D
prior.D$P <- P
res[[1]] <- TIMBR(corr[, trait], prior.D, prior.M1)
```

Repeat for QTL at 178.2520 Mb.

```{r TIMBR_2, eval=FALSE}
b2 <- eff.df[9, 6]
b3 <- eff.df[10, 6]
corr <- pheno.raw
corr[addcovar$block2 == 1, "IL10"] <- corr[addcovar$block2 == 1, "IL10"] - b2
corr[addcovar$block3 == 1, "IL10"] <- corr[addcovar$block3 == 1, "IL10"] - b3
P <- model.probs.reduced[, , probes[2]]
prior.D <- mcv.data$prior.D
prior.D$P <- P
res[[2]] <- TIMBR(corr[, trait], prior.D, prior.M1)
save(res, file="TIMBR_res_IL10_fit_coef.Rdata")
```

Plot results

```{r plots, fig.align='center'}
load("TIMBR_res_IL10_fit_coef.Rdata")
for (i in 1:2) {
  cat("IL10 @ ", pos[i], " Mb\n")
  TIMBR::TIMBR.plot.haplotypes(res[[i]])
  TIMBR::TIMBR.plot.circos(res[[i]])
}
```

##### Conclusions

1. IL10 mapping should include the batch covariates. In any event it would be hard to justify NOT using batch covariates.

2. The TIMBR runs with the appropriately corrected phenotype values yield results very similar to those obtained with the original batch-corrected values. Why including the covariates in the TIMBR runs gave spurious results is a mystery. Probably my error in constructing the "Z" design matrix.
