---
title: "Trivariate mapping of cluster traits"
author: "Frederick J. Boehm"
date: "3/23/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Read the csv files

```{r}
cluster_traits <- readr::read_csv("../data/pheno679_eigenscores_from_WGCNA_unsigned.csv")
peaks <- readr::read_csv("../data/pheno679_eigenscores_unsigned_maxLOD_peaks.csv")
```
```{r}
load("../data/reduced_map_qtl2_mapping_objects.Rdata")
```

```{r}
phe <- as.matrix(cluster_traits[, c(2, 8, 18)])
rownames(phe) <- cluster_traits$X1
start <- max(which(map.1$`10` < 20)) 
end <- min(which(map.1$`10` > 45))
fn <- "../data/cluster-trivariate-mapping.rds"
if (!file.exists(fn)){
  sp_out <- qtl2pleio::scan_pvl(probs = genoprobs.1$`10`, 
                                pheno = phe,
                                addcovar = addcovar, 
                                start_snp = start,
                                n_snp = end - start + 1, 
                                n_cores = parallel::detectCores() - 1
                                )
saveRDS(sp_out, file = fn)
} else {
  sp_out <- readRDS(fn)
}
```

```{r}
library(magrittr)
sp_out %>%
  qtl2pleio::calc_lrt_tib()
```
```{r}
sp_out %>%
  qtl2pleio::calc_profile_lods() %>%
  qtl2pleio::add_pmap(pmap = map.1$`10`) %>%
  ggplot2::ggplot() + ggplot2::geom_line(ggplot2::aes(x = marker_position, y = profile_lod, colour = trait))
```
## Pairwise analyses

First, the analyses for traits for the first two columns of `phe`, which correspond to columns 2 and 8 of `cluster_traits`. 

```{r}
sp_out <- qtl2pleio::scan_pvl(probs = genoprobs.1$`10`, 
                                pheno = phe[, 1:2],
                                addcovar = addcovar, 
                                start_snp = start,
                                n_snp = end - start + 1, 
                                n_cores = parallel::detectCores() - 1
                                )
(lrt <- qtl2pleio::calc_lrt_tib(sp_out))
```

```{r}
sp_out %>%
  qtl2pleio::calc_profile_lods() %>%
  qtl2pleio::add_pmap(pmap = map.1$`10`) %>%
  ggplot2::ggplot() + ggplot2::geom_line(ggplot2::aes(x = marker_position, y = profile_lod, colour = trait))
```

```{r}
pp_index <- qtl2pleio::find_pleio_peak_tib(tib = sp_out, 
                                           start_snp = start
                                           )
b_out <- qtl2pleio::boot_pvl(probs = genoprobs.1$`10`, 
                    pheno = phe[, 1:2],
                    addcovar = addcovar, 
                    start_snp = start,
                    n_snp = end - start + 1, 
                    n_cores = parallel::detectCores() - 1, 
                    pleio_peak_index = pp_index,
                    nboot_per_job = 400
                    )
(pvalue <- mean(b_out >= lrt))
```


Now, the pairwise analyses for traits 1 and 3 in `phe`, which correspond 
to columns 2 and 18 in `cluster_traits`.

```{r}
sp_out <- qtl2pleio::scan_pvl(probs = genoprobs.1$`10`, 
                                pheno = phe[, c(1,3)],
                                addcovar = addcovar, 
                                start_snp = start,
                                n_snp = end - start + 1, 
                                n_cores = parallel::detectCores() - 1
                                )
(lrt <- qtl2pleio::calc_lrt_tib(sp_out))
```






```{r}
sp_out %>%
  qtl2pleio::calc_profile_lods() %>%
  qtl2pleio::add_pmap(pmap = map.1$`10`) %>%
  ggplot2::ggplot() + ggplot2::geom_line(ggplot2::aes(x = marker_position, y = profile_lod, colour = trait))
```

```{r}
pp_index <- qtl2pleio::find_pleio_peak_tib(tib = sp_out, 
                                           start_snp = start
                                           )
b_out <- qtl2pleio::boot_pvl(probs = genoprobs.1$`10`, 
                    pheno = phe[, c(1, 3)],
                    addcovar = addcovar, 
                    start_snp = start,
                    n_snp = end - start + 1, 
                    n_cores = parallel::detectCores() - 1, 
                    pleio_peak_index = pp_index,
                    nboot_per_job = 400
                    )
(pvalue <- mean(b_out >= lrt))
```




Now, the pairwise analyses for traits 2 and 3 in `phe`, which correspond 
to columns 8 and 18 in `cluster_traits`.



```{r}
sp_out <- qtl2pleio::scan_pvl(probs = genoprobs.1$`10`, 
                                pheno = phe[, 2:3],
                                addcovar = addcovar, 
                                start_snp = start,
                                n_snp = end - start + 1, 
                                n_cores = parallel::detectCores() - 1
                                )
(lrt <- qtl2pleio::calc_lrt_tib(sp_out))
```

```{r}
sp_out %>%
  qtl2pleio::calc_profile_lods() %>%
  qtl2pleio::add_pmap(pmap = map.1$`10`) %>%
  ggplot2::ggplot() + ggplot2::geom_line(ggplot2::aes(x = marker_position, y = profile_lod, colour = trait))
```

```{r}
pp_index <- qtl2pleio::find_pleio_peak_tib(tib = sp_out, 
                                           start_snp = start
                                           )
b_out <- qtl2pleio::boot_pvl(probs = genoprobs.1$`10`, 
                    pheno = phe[, 2:3],
                    addcovar = addcovar, 
                    start_snp = start,
                    n_snp = end - start + 1, 
                    n_cores = parallel::detectCores() - 1, 
                    pleio_peak_index = pp_index,
                    nboot_per_job = 400
                    )
(pvalue <- mean(b_out >= lrt))
```

```{r}
colnames(cluster_traits)[c(2, 8, 18)]
```

### Colophon

This report was generated on `r Sys.time()` using the following computational environment and dependencies: 

```{r colophon, cache = FALSE}
# which R packages and versions?
if ("devtools" %in% installed.packages()) devtools::session_info()
```

The current Git commit details are:

```{r}
# what commit is this file at? 
if ("git2r" %in% installed.packages() & git2r::in_repository(path = ".")) git2r::repository(here::here())  
```


