---
title: "Clinical traits: rescans after genotype collapsing to CAST v not CAST"
author: "Frederick J. Boehm"
date: "11/14/2019"
output: github_document
params:
  lastmod: !r lubridate::now()
---

## Goals

The goal here is to rescan clinical traits after collapsing genotypes to CAST v not CAST.

IL-9 has a peak that shows CAST v not CAST allele partitioning.

We omit batch covariates because we work with batch-adjusted clinical traits.

We treat mice from a single line as genetically identical.


```{r}
library(tidyverse)
```

```{r read-traits}
traits <- readr::read_csv("../data/batch_corrected_phenos_individual_05OCT2019_v4.csv")

```

```{r load-genotypes}
probs_line <- readRDS("../data/genotypes_array.rds")
```

We create a probs object that assumes that each mouse from a shared line is genetically identical.

```{r per-mouse-genotypes}
# first, how many mice per line are there??
table(traits$CC.Line) -> tab
(tib <- tibble::tibble(line = names(tab), count = tab))
```

```{r rownames}
rn_pre <- purrr::map2(.x = tib$line, 
                  .y = tib$count, 
                  .f = function(x, y){rep(x, y)})
rn <- rn_pre %>% unlist()
```

Now, make a 162 by 36 by n.markers array from the 60 by 36 by n.markers

```{r make_probs_array}
probs <- array(dim = c(162, 36, ))
```


## Apply to all elements of `results` list

```{r, scan1_loop}
s1perm <- list()
s1perm_collapsed <- list()
s1_collapsed <- list()
s1 <- list()
for (i in seq_along(1:length(results))){
  r1 <- results[[i]]
  tib <- tibble(partition = names(table(r1$post.M)),
                count = table(r1$post.M)) %>%
    arrange(desc(count))
  part <- tib$partition[1] %>% # choose the top row in tib
    stringr::str_split(pattern = ",") %>%
    unlist() %>%
    as.numeric()
  gg <- qtl2effects::collapse_geno(genoprobs.1[[peaks$chr[i]]], part)
  dimnames(gg)[[2]] <- as.character(0:max(part))
  dimnames(gg)[[3]] <- names(map.1[[peaks$chr[i]]])
  gg_named <- list()
  gg_named$`1` <- gg
  attributes(gg_named)$class <- c("calc_genoprob", "list")
  ## make trait
  trait_name <- peaks$lodcolumn[i]
  trait <- traits %>%
    select(trait_name) %>%
    as.matrix() %>%
    (function(x) {rownames(x) <- traits$mouse_id; 
    colnames(x) <- trait_name; 
    return(x)}
    )
  ## scan1 calls
  s1_collapsed[[i]] <- qtl2::scan1(genoprobs = gg_named, 
              pheno = trait, 
              addcovar = addcovar
              )
  gg_full <- list()
  gg_full$`1` <- genoprobs.1[[peaks$chr[i]]]
  attributes(gg_full)$class <- c("calc_genoprob", "list")
  s1[[i]] <- qtl2::scan1(genoprobs = gg_full, 
              pheno = trait, 
              addcovar = addcovar
              )
  ## scan1perm calls
  s1perm_collapsed[[i]] <- qtl2::scan1perm(genoprobs = gg_named, 
              pheno = trait, 
              addcovar = addcovar, 
              n_perm = 10000
              )
  #s1perm[[i]] <- qtl2::scan1perm(genoprobs = genoprobs.1[[peaks$chr[i]]], 
  #            pheno = trait, 
  #            addcovar = addcovar, 
  #            n_perm = 10000
  #            )
}

```

## Calculate quantiles for each set of permutations

```{r get-quantiles}
(qmat <- sapply(X = s1perm_collapsed, FUN = quantile, probs = c(0.95, 0.99, 0.999, 0.9999)))
```

```{r collapsed-lods}
tibble(trait = names(sapply(X = s1_collapsed, FUN = max)), 
       collapsed_lod = sapply(X = s1_collapsed, FUN = max)) %>%
  bind_cols(as_tibble(t(qmat)))
```
