---
title: "Neto hotspot 4 analysis per Rick's email"
author: "Frederick J. Boehm"
date: "11/18/2019"
output: github_document
params:
  lastmod: !r lubridate::now()
  hot: 4
---

Last modified: `r params$lastmod`.

## Text of Rick's email on Nov 18, 2019

> Sorry for the misunderstanding. I’m only interested in the hotspot4 traits identified in the analysis I sent on Friday afternoon (hotspots_allele_effects2.html). There are 11 traits. The trait/probe information was in my Email message, but I add it here again.  /rb

> ## 4:Chr10.34 
> ## RVBD_0158 :  JAX00302006 
> ## RVBD_0437c :  JAX00301562 
> ## RVBD_0490 :  UNC18950608 
> ## RVBD_1051c :  UNC18908602 
> ## RVBD_1493 :  UNC18891985 
> ## RVBD_1692 :  UNC18962410 
> ## RVBD_2643A :  UNC18925693 
> ## RVBD_2854 :  JAX00300565 
> ## RVBD_3261 :  UNCHS029542 
> ## RVBD_3262 :  UNCHS029517 
> ## RVBD_3452 :  JAX00302076



## Analysis

```{r pkgs}
library(dplyr)
library(TIMBR)
```



```{r read-genotypes}
probs <- readRDS("../data/genotypes_array.rds")
traits <- readRDS("../data/tnseq-traits.rds")
```

```{r read-csv}
tt <- read.csv("../data/neto_traits_by_probe3_annotated.csv")
neto <- tt %>%
  tidyr::pivot_longer(cols = V2:V35, values_to = "trait", names_to = "trait_name") %>%
  dplyr::filter(!is.na(trait)) %>%
  dplyr::select(- trait_name, - neto.n, - row, - n.traits, - cM, -hs, - chr)
neto_plus <- tt %>%
  tidyr::pivot_longer(cols = V2:V35, values_to = "trait", names_to = "trait_name") %>%
  dplyr::filter(!is.na(trait)) %>%
  dplyr::select(- trait_name)

traits_timbr_annotated <- neto_plus %>%
  dplyr::select(probe, trait) %>%
  purrr::pmap( 
           .f = function(probe, trait){
             pheno <- traits[ , colnames(traits) == trait, drop = FALSE]
             geno <- probs[ , , dimnames(probs)[[3]] == probe]
             qtl2::fit1(genoprobs = geno, 
                        pheno = pheno, 
                        )
           }
             ) %>%
  purrr::map(.f = function(x){
    tibble::tibble(lod = x$lod)
  }) %>%
  bind_rows() %>%
  bind_cols(neto_plus) %>%
  dplyr::select(probe, trait, lod, hs) %>%
  dplyr::group_by(hs) %>%
  dplyr::group_by(trait) %>%
  dplyr::filter(lod == max(lod)) %>%
  dplyr::ungroup() %>%
  dplyr::ungroup()
```

```{r, annotation-tibble}
(hot_indices <- neto_plus %>%
  dplyr::group_by(hs) %>%
  dplyr::tally() %>%
  dplyr::mutate(end = cumsum(n)) %>%
  dplyr::mutate(start = 1L + end - n)
)
```

### TIMBR setup

```{r prior_M-define}
##### From GK example code
# Specify allelic series prior
# Suggested by Wes
# Influences how much prior weight it places on more or less complicated allelic series
prior_M <- list(model.type = "crp", # crp - Chinese Restaurant Process
                prior.alpha.type = "gamma",
                prior.alpha.shape = 1,
                prior.alpha.rate = 2.333415)
```


```{r load-data-timbr}
data(mcv.data) # get A matrix
```

```{r make_neto_list}
neto_list <- tibble(probe = c("JAX00302006",
                    "JAX00301562", 
                    "UNC18950608", 
                    "UNC18908602",
                    "UNC18891985",
                    "UNC18962410",
                    "UNC18925693", 
                    "JAX00300565", 
                    "UNCHS029542",
                    "UNCHS029517", 
                    "JAX00302076"),
                    trait = c("RVBD_0158",
                              "RVBD_0437c", 
                              "RVBD_0490", 
                              "RVBD_1051c", 
                              "RVBD_1493",
                              "RVBD_1692",
                              "RVBD_2643A",
                              "RVBD_2854",
                              "RVBD_3261",
                              "RVBD_3262",
                              "RVBD_3452"
                              )
                    ) %>% 
  apply(FUN = as.list, MARGIN = 1)

```

### TIMBR call

```{r }
outfn <- "../data/timbr-rick-11-traits.rds"
# ensure that inputs to call_timbr all have subjects in same order!
subject_ids <- rownames(traits)
##
indices_gp <- match(subject_ids, rownames(probs))
gp <- probs[indices_gp, , ]
##
if (!file.exists(outfn)){
  timbr_out <- parallel::mclapply(neto_list, 
  #timbr_out <- lapply(neto_small, 
                                  FUN = qtl2effects::call_timbr, 
                                  mc.cores = parallel::detectCores(),
                                  traits_df = traits,
                                  prior_M = prior_M, 
                                  genoprobs_array = gp,
                                  addcovar = NULL
                                  )
  saveRDS(timbr_out, outfn)
} else {
  timbr_out <- readRDS(outfn)
}
```

## Examining TIMBR results

We have the object `timbr_out`, a list of length `r length(timbr_out)`. Each element of `timbr_out` corresponds to a single TIMBR analysis, and is itself a list. The names of the outputted objects (for a single TIMBR analysis) are:

```{r names}
names(timbr_out[[1]])
```

We can access the 10,000 allelic series (per analysis) by its name `post.M`, as in the following code:

```{r get-postM}
tr1_postM <- timbr_out[[1]]$post.M
```

## TIMBR density plots

```{r plots}
lapply(X = timbr_out, 
       FUN = TIMBR::TIMBR.plot.haplotypes
       )
```

