---
title: "QTL mapping for Tnseq WCGNA traits"
output: 
  html_document: 
    code_folding: hide
params:
  lastmod: !r lubridate::now()
---

```{r setup2, include=FALSE, dev="CairoPNG"}
knitr::opts_chunk$set(dev="CairoPNG")
```



Last knitted: `r params$lastmod`.



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
nc <- 32
```

## Overview

We want to find the QTL peaks for the WGCNA traits that Rick Baker determined. 

Once we have those peaks, we'll use TIMBR at each peak to get the allelic series.

We'll also look at genotype probabilities (at QTL peak position) per mouse line.

We'll look at the patterns of genotype probabilities among mouse lines in each cluster, where the clusters are defined as in Clare's preprint. In short, there are six clusters and 
we want to see if there are distinct immunological responses for the different clusters.

## Step 1: map the Tnseq WGCNA traits

```{r}
library(magrittr)
```


```{r}
load("../data/reduced_map_qtl2_mapping_objects.Rdata")
pheno <- readr::read_csv("../data/pheno_eigenscores_from_WGCNA_all60_win.csv") %>%
  dplyr::rename(mouse_line = X1)

pheno_matrix <- as.matrix(pheno[, -1])
rownames(pheno_matrix) <- pheno$mouse_line
```

```{r}
kinship <- qtl2::calc_kinship(probs = genoprobs.1, type = "loco", cores = nc)
s1 <- qtl2::scan1(genoprobs = genoprobs.1, pheno = pheno_mat, kinship = kinship, reml = TRUE, cores = nc)
```

```{r}
peaks <- qtl2::find_peaks(scan1_output = s1, map = map.1, peakdrop = 1.5, prob = 0.95, cores = nc, threshold = 6) %>%
  tibble::as_tibble()
```

## Step 2: TIMBR at QTL peak positions




```{r}
gp <- readRDS("../data/cc36.rds")
```
```{r TIMBR-setup}
mapdf <- qtl2convert::map_list_to_df(map.1)
fn <- "../data/cc-tnseq-modules-timbr.rds"
if (!file.exists(fn)){
  prior_M1 <- list(model.type = "crp", # crp - Chinese Restaurant Process
                  prior.alpha.type = "gamma",
                  prior.alpha.shape = 1,
                  prior.alpha.rate = 2.333415)
  prior_M2 <- list(model.type = "crp",
                  prior.alpha.type = "gamma",
                  prior.alpha.shape = 2.3009322,
                  prior.alpha.rate = 0.7488104
                  )
  library(TIMBR)
  data(mcv.data) # get A matrix
  set.seed(3411192) # to ensure getting the same samples with TIMBR
  t_out_peaks <- peaks %>%
    dplyr::mutate(marker_index = purrr::map2_int(.x = pos, .y = chr, .f = function(x, y){
      pp <- mapdf %>%
        dplyr::filter(chr == y, x == pos) %>%
        dplyr::select(pos) %>%
        unlist()
      # some markers have identical map positions
      return(which(map.1[[y]] == pp[1])[1]) 
      })) %>%
    dplyr::filter(chr != "X") %>% # No TIMBR for X chr yet!
    dplyr::filter(lod >= 6) %>%
    dplyr::mutate(timbr = purrr::pmap(.l = list(lodcolumn, chr, marker_index), 
                                      .f = function(lodcolumn, chr, marker_index){
                                        qtl2tb::run_timbr(lodcolumn, 
                                                          chr, 
                                                          marker_index, 
                                                          gp,
                                                          pheno = pheno_matrix[, lodcolumn, drop = FALSE],
                                                          addcovar = NULL)
                                      }
                          ))
  rm(gp)
  saveRDS(t_out_peaks, fn)
} else {
  t_out_peaks <- readRDS(fn)
}
```

I need to write a function that pulls the genotype probabilities object for each
QTL peak position. 


