---
title: "QTL mapping for Tnseq WCGNA traits and CFU and Cytokines"
output: 
  html_document: 
    code_folding: hide
params:
  lastmod: !r clock::zoned_now("America/New_York")
---

```{r setup2, include=FALSE, dev="CairoPNG"}
knitr::opts_chunk$set(dev="CairoPNG")
```



Last knitted: `r params$lastmod`.



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
#nc <- 32
nc <- 2
```

## Overview

This is the analysis that Chris requested.

We want to find the QTL peaks for the WGCNA traits that Rick Baker determined. 

Once we have those peaks, we'll use TIMBR at each peak to get the allelic series.

We'll also look at founder allele dosages (at QTL peak position) per mouse line.

We'll look at the patterns of genotype probabilities among mouse lines in each cluster, where the clusters are defined as in Clare's preprint. In short, there are six clusters and 
we want to see if there are distinct immunological responses for the different clusters.

## Step 1: map the Tnseq WGCNA traits

```{r}
library(magrittr)
```


```{r}
load("../data/reduced_map_qtl2_mapping_objects.Rdata")
pheno <- readr::read_csv("../data/pheno_eigenscores_from_WGCNA_all60_win.csv") %>%
  dplyr::rename(mouse_line = X1)

pheno_matrix <- as.matrix(pheno[, -1])
rownames(pheno_matrix) <- pheno$mouse_line
```

```{r}
kinship <- qtl2::calc_kinship(probs = genoprobs.1, type = "loco", cores = nc)
s1 <- qtl2::scan1(genoprobs = genoprobs.1, pheno = pheno_matrix, kinship = kinship, reml = TRUE, cores = nc)
```

```{r}
fn <- "../data/cc-tnseq-modules-timbr.rds"
```

```{r, child = "tnseq-wgcna-mapping-peaks.Rmd"}
```



## CFU and cytokines analysis


```{r}
pheno <- readr::read_csv("../data/pheno_clinical_4qtl2_win2_05OCT2019_v4.csv") %>%
  dplyr::rename(mouse_line = X1)

pheno_matrix <- as.matrix(pheno[, -1])
rownames(pheno_matrix) <- pheno$mouse_line
```

```{r}
s1 <- qtl2::scan1(genoprobs = genoprobs.1, pheno = pheno_matrix, kinship = kinship, reml = TRUE, cores = nc)
fn <- "../data/cc-clinical-timbr.rds"
```

```{r, child = "tnseq-wgcna-mapping-peaks.Rmd"}
```

